\chapter*{Acknowledgments}

Embarking on the arduous yet exhilarating journey of a PhD, one finds himself traversing an emotional rollercoaster of intellectual growth, self-discovery, and relentless perseverance. Approaching its conclusion, I realise now how well-surrounded I was throughout these four years I spent in the SerVal team of the SnT research centre. Here, I want to say a few words to those who took an important part in this adventure and had positive impacts on it.

Yves, these first words go for you. As my official supervisor, you trusted me since the very first time we met and you gave me all the keys I needed to successfully pursue my PhD. I am thankful for your support and your constant positive thinking.

Mike, you were my direct supervisor and when I joined the team I didn't fully measure its excellence. You are largely contributing to it and I learned so much from you. I am grateful for all your counselling, help and every one of the discussions we had. I admire the amount of work you give to create this wonderful research environment.

Sarra, our paths met at a crucial point in my life. I was not expecting your arrival in our team nor that I would benefit from the additional support and expertise of a young researcher. I am deeply thankful for your guidance and cooperation. You genuinely inspired me and will continue to do so for a long time.

Maxime, while you were expanding the domains of expertise in our team, you always remained available and supportive to me. I sincerely appreciate your dedication and I will keep good memories of the discussions we had.

I would like to thank all my co-authors for their involvement and feedback in the different projects. Thanks also to the jury members for the time and interest they gave to my research. This dissertation would not have been the same without your help. I also want to express my gratitude towards all my fellow SerVal colleagues, some arrived, and some left through my journey but all contributed to creating a fantastic place to conduct research. I hope we will stay in touch.

On a more personal note, I wish to thank my family for their support: my mom, who listened so many times to my work-related stories, my dad, who doesn't know how important he is to me and my brother, whose presence has the magical ability to always transport me back to the carefree days of childhood. You are the joy of my life. Finally, I thank all my friends with whom I was able to temper the most stressful part of this adventure.